Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-06-23 paper.qmd
Update with xbins 12

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 10, 
  r2 = r2,
  q = 0.05)

@@

```{r}
#| echo: false
#| label: fig-bins-scurve
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 91$ (7, 13), (b) $b = 190$ (10, 19), and (c) $b = 364$ (14, 26). The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin. What is the number of bins that would be effective?"
#| fig-width: 6
#| fig-height: 3

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 190, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, 
            angle=90),
        title = element_text(size = 36))
```

@@

scurve_model <- fit_highd_model(
  training_data = training_data_scurve,
  emb_df = umap_scurve_scaled,
  bin1 = effective_bin1_scurve,
  r2 = r2,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

@@

<!--Using first quantile as the benchmark value-->
```{r}
hex_grid_with_low_counts <- full_join(hex_grid2, 
    df_bin_centroids_scurve, 
    by = c("hex_poly_id" = "hexID")) |>
  filter(!is.na(std_counts)) |>
  mutate(type = if_else(std_counts <= benchmark1, "low", "high")) |>
  filter(type == "low")

>>>>

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.07)

@@

```{r}
#| echo: false
#| label: fig-bins-scurve
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 91$ (7, 13), (b) $b = 264$ (12, 22), and (c) $b = 364$ (14, 26). The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin. What is the number of bins that would be effective?"
#| fig-width: 6
#| fig-height: 3

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 264, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, 
            angle=90),
        title = element_text(size = 36))
```

@@

@@

<!--Using first quantile as the benchmark value-->
```{r}
hex_grid_with_counts_s_curve2 <- full_join(
  hex_grid2, 
  df_bin_centroids_scurve |> select(hexID, std_counts), 
  by = c("hex_poly_id" = "hexID"))

hex_grid_with_low_counts <- full_join(hex_grid2, 
    df_bin_centroids_scurve, 
    by = c("hex_poly_id" = "hexID")) |>
  filter(!is.na(std_counts)) |>
  mutate(type = if_else(std_counts <= benchmark1, "low", "high")) |>
  filter(type == "low")
